### 2020-01-10 work.Rmd
dif_intuitive works great for plotting logit graphs

```{r}
library(tidyverse)
library(mirt)
R.utils::sourceDirectory("R")
```

```{r}
pars_1_dif <-
    read_csv("data-parameters/kopf2015-8.csv") %>%
    arrange(desc(easy)) %>%
    mutate(
        item = row_number(),
        nuisance_disc = c(0, 0, 0, 0.5, 0, 0, 0, 0)
    )

sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)
```

```{r}
dif_constrained_baseline_one_at_a_time(data = sim_1_dif$data, groups = sim_1_dif$groups)
```

```{r}
mod <- 
    multipleGroup(
        sim_1_dif$data,
        1,
        itemtype = "Rasch",
        sim_1_dif$groups,
        invariance = c("free_var"),
        technical = list(NCYCLES = 100),
        SE = TRUE
    )

mod %>% 
    mod_to_draws_df() %>% 
    draws_df_to_logit_plot()
```

```{r}
pars_3_dif <-
    read_csv("data-parameters/kopf2015-8.csv") %>%
    arrange(desc(easy)) %>%
    mutate(
        item = row_number(),
        nuisance_disc = c(0, 0, 0, 0.5, 0.5, 0.5, 0, 0)
    )
```