### 2020-01-13 R-work/work.Rmd
work.Rmd works great for summarizing the all other method!

```{r}
library(tidyverse)
library(mirt)
R.utils::sourceDirectory("R")
```

```{r}
pars_1_dif <-
    read_csv("data-parameters/kopf2015-8.csv") %>%
    arrange(desc(easy)) %>%
    mutate(
        item = row_number(),
        nuisance_disc = c(0, 0, 0, 0.1, 0.1, 0, 0, 0)
    )

sim_1_dif <- sim_bias(pars = pars_1_dif, n_ref = 50000, n_foc = 50000, mean_foc = -1)

sim <- read_rds("data-documentation/sim.rds")

sim %>% mod_to_draws_df() %>% draws_df_to_logit_plot()
```

```{r}
all_other <- dif_all_other(sim_1_dif$data, sim_1_dif$groups)

iterative_backward <- dif_iterative_backward(sim_1_dif$data, sim_1_dif$groups, all_other)
```