### 2020-01-13 paper/paper1.Rmd

# Hello

but there is no guarantee that it will work for the small sample sizes with which psychometricians often work [@mcdonald1995goodness].