### 2020-01-13 work.Rmd

```{r}
library(tidyverse)
library(mirt)
set.seed(123)
R.utils::sourceDirectory("R")
```

```{r}
create_pars_1_biased_item <- function(){
    tibble(
        item = 1:8,
        easy = rnorm(8, 0, 1),
        target_disc = 1,
        nuisance_disc = c(rep(0.5, 1), rep(0, 7))
    ) %>%
    arrange(desc(nuisance_disc), easy)
}
```

```{r}
out <- 
    tibble(
        pars = rerun(1, create_pars_1_biased_item())
    ) %>% 
    mutate(
        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
    ) %>% 
    mutate(
        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
        all_other_mod = 
            map2(
                sim, 
                all_other_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex")
                )
            ),
        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
    )

out$intuitive_mod[[1]] %>% 
    mod_intuitive_to_draws_df() %>% 
    draws_df_to_logit_plot() +
    geom_point(data = out$all_other_final_dif[[1]], aes(x = dif, y = item), color = "red") +
    geom_vline(xintercept = filter(out$all_other_final_dif[[1]], anchor)$dif[1], linetype = "dashed", color = "red")
```