### 2020-01-16 paper/paper1.Rmd
great day of progress

# Intro

Inspired by @camilli1992conceptual, we view differential item functioning (DIF) as bias towards a group of students that manifests when we impose an item response model with too few ability dimensions. From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item. For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability but performance also depends on their verbal ability. In this case, math ability is the "target ability" and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test will result in students with low verbal ability being given a score systematically lower than their true math ability; this is DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, @kopf2015framework simulate students as belonging to either the reference or the focal group. The item difficulties for the reference group, ${b_j}^{\text{ref}}$, are fixed. The item difficulties for the focal group are ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits. Student ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(0,1)$ for students in the focal group. The item responses are generated according to the Rasch model. The Rasch model specifies that the probability of student $i$ responding correctly to item $j$ is 
$$
P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i - b_j)
$$
where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.

We find it more informative to describe simulation conditions in terms of single dimensional items and multidimensional abilities. For example, the @kopf2015framework simulation study can be translated from the DIF-as-item-property view to the DIF-as-student-property view. Item difficulty is set to what was previously ${b_j}^{\text{ref}}$ for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(0,1)$ for students in the focal group. The nuisance ability is set to $\eta_i^{\text{ref}} = 0$ for the reference group and $\eta_i^{\text{ref}} = -1$ for the focal group. We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF and $a_{2j} = 0$ otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional 2PL model, the probability student $i$ responding correctly to item $j$ is

$$
P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i - b_j).
$$
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which the variance of the nuisance ability is set to 0 for all students.

If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower-than-necessary dimensional model? @camilli1992conceptual tested this idea with the goal of a "more satisfying description of the secondary abilities." They found that the rotational indetermincy of item response models is difficult to overcome and concluded that "a priori knowledge of the true factor structure" is necessary. It's hard to imagine how one would have such knowledge. Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias that manifests in the item parameters.

Psychometricians have long been in search of the perfect DIF detection method. @meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA) method where each item is independently tested for DIF using all other items as anchors. More specifically, testing is done using a nested model comparison between the baseline model where all item parameters are fixed across groups and the flexible model where the parameters of the tested item are freed across groups. For example, @edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. One might wonder how we can be sure its those 10 items and not the other 11 items with DIF. Implicit in the use of AOAA is the assumption. Most applications of DIF and many DIF simulation studies do not make explicit the assumptions of the DIF methods. STroble 2018 lists some of these assumptions. For AOAA the assumption is that. It's easy to show that AOAA is imperfect: Imagine a test with 3 items. Each item will test positive for DIF.  a phenomenon that went indirectly address in the literature until andrich at all named it artificial DIF.

We advocate for methods that present information to the analyst where the analyst makes decisions as opposed to methods which automate the detection and subsequently the removal of items with DIF as is frequently the case (e.g. in PISA). The most obvious method is to at first punt on disentangling between ability and bias and instead set both group means to zero and estimate the item parameters separately for each group. The difference in item parameters, then, includes both the difference in ability and bias. To our knowledge, this method has not been suggested in the literature with the exception of pohl who uses it in a pedagogical example and tashjkdf who uses it pre to post. For example, consider a didactic scenario in which an eight item exam where the focal group and reference group. Where the difference in ability between groups is one logit. And additionally, the first two items have a one logit bias against the students who. We fit a model then fix the means. We then sample as leah describes (look at how she describes it) then we get this plot. It cannot be said strongly enough that this plot contains all of the information about the combination of ability and bias that is available in the data. It is then up to the analyst and/or method to choos what's known anchor point, which is essentially the difference in abilities with remaining differences being attributed to bias.