### 2020-02-24 R-work-in-progress/work-rough.Rmd
its working and just doing gini stuff now

- implement gini!!

- implement the forward SA

```{r}
library(tidyverse)
library(mirt)
library(cluster)
set.seed(123)
R.utils::sourceDirectory("R")
```

# 1

```{r}
out_1_biased_item <- 
    tibble(
        pars = rerun(100, create_pars_1_biased_item()) # swap in pars function here
    ) %>% 
    # sim and fit intuitive
    mutate(
        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
    ) %>% 
    # dif: all other
    mutate(
        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
        all_other_mod = 
            map2(
                sim, 
                all_other_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
    ) %>% 
    # dif: iterative backward
    mutate(
        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
        it_back_mod = 
            map2(
                sim, 
                it_back_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
    ) %>% 
    # dif: tibs cluster
    mutate(
        cluster = 
            intuitive_mod %>% 
            map(mod_to_final_dif) %>% 
            map(~ mutate(., cluster = irt_cluster(dif)))
    ) %>% 
    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>% 
    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>% 
    mutate(
        cluster_mod = 
            map2(
                sim, 
                cluster_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
    )
```

# 3

```{r}
out_3_biased_items <- 
    tibble(
        pars = rerun(1, create_pars_3_biased_items())
    ) %>% 
    # sim and fit intuitive
    mutate(
        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
    ) %>% 
    # dif: all other
    mutate(
        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
        all_other_mod = 
            map2(
                sim, 
                all_other_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
    ) %>% 
    # dif: iterative backward
    mutate(
        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
        it_back_mod = 
            map2(
                sim, 
                it_back_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
    ) %>% 
    # dif: tibs cluster
    mutate(
        cluster = 
            intuitive_mod %>% 
            map(mod_to_final_dif) %>% 
            map(~ mutate(., cluster = irt_cluster(dif)))
    ) %>% 
    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>% 
    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>% 
    mutate(
        cluster_mod = 
            map2(
                sim, 
                cluster_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
    )
```

# 4

```{r}
out_4_biased_items_balanced <- 
    tibble(
        pars = rerun(100, create_pars_4_biased_items_balanced()) # swap in pars function here
    ) %>% 
    # sim and fit intuitive
    mutate(
        sim = pars %>% map(~ sim_bias(pars = ., n_ref = 10000, n_foc = 10000, mean_foc = -1)),
        intuitive_mod = sim %>% map(~ fit_mod_intuitive(.$data, .$groups))
    ) %>% 
    # dif: all other
    mutate(
        all_other_status = sim %>% map(~ dif_all_other(.$data, .$groups)),
        all_other_mod = 
            map2(
                sim, 
                all_other_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        all_other_final_dif = all_other_mod %>% map(mod_to_final_dif)
    ) %>% 
    # dif: iterative backward
    mutate(
        it_back_status = map2(sim, all_other_status, ~ dif_iterative_backward(.x$data, .x$groups, .y)),
        it_back_mod = 
            map2(
                sim, 
                it_back_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        it_back_final_dif = it_back_mod %>% map(mod_to_final_dif)
    ) %>% 
    # dif: tibs cluster
    mutate(
        cluster = 
            intuitive_mod %>% 
            map(mod_to_final_dif) %>% 
            map(~ mutate(., cluster = irt_cluster(dif)))
    ) %>% 
    mutate(cluster_stats = cluster %>% map(get_cluster_stats)) %>% 
    mutate(cluster_status = map2(cluster, cluster_stats, left_join, by = "cluster")) %>% 
    mutate(
        cluster_mod = 
            map2(
                sim, 
                cluster_status, 
                ~ mod_flexible(
                    .x$data, 
                    .x$groups, 
                    which(.y$status == "flex"),
                    se = TRUE
                )
            ),
        cluster_final_dif = cluster_mod %>% map(mod_to_final_dif)
    )
```

```{r}
save.image("work_run_1.Rdata")
```