### 2020-02-24 paper/TEST.Rmd

---
output:
  bookdown::pdf_document2:
    includes:
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 2
header-includes: 
- \usepackage{float}
- \usepackage[autostyle, english = american]{csquotes}
- \setlength{\parindent}{4em}
- \setlength{\parskip}{2em}
- \usepackage{setspace}\doublespacing
- \usepackage{amsmath}
- \DeclareMathOperator*{\argmax}{arg\,max}
fontsize: 11pt
link-citations: yes
title: "Measurement invariance"
subtitle: |
    Ben Stenhaug & Ben Domingue  
    Stanford University
abstract: |
  ABSTRACT TO DO
  <!-- https://tinyurl.com/ybremelq -->
bibliography: paper.bib
---

```{r setup, include = FALSE}
# 
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

knitr::opts_chunk$set(
    fig.align='center',
    echo=FALSE,
    warning=TRUE,
    fig.pos='H',
    fig.retina = 5
)

library(tidyverse)
library(ggtext)
library(here)
library(mirt)
library(cluster)
library(DescTools)
library(knitr)
library(kableExtra)
set.seed(123)
R.utils::sourceDirectory(here("R"))

theme_set(theme_bw())
```

The Gini index is maximized to select the anchor point. The intuition is that we want to find the anchor point such that a small minority of items have bias while the vast majority do not. Denote the easiness difference between the two groups when both group means are fixed to zero as $d_j = {b_j}^\text{ref} - {b_j}^\text{foc}$, a function that calculates the Gini index from a vector of non-negative elements as $G(\mathbf{x})$, and the mean of the focal group ability when the reference group ability is fixed to zero (i.e. the anchor point) as $\Theta$.

$$
\argmax_\Theta G(\mathbf{d} + \Theta)
$$