### 2020-02-24 paper/paper_3_uncertainty.Rmd

We know that this test contains DIF. Which items depends on ability difference which we can be sure of. We can, however, quantify the amount of DIF on this test. We can take the SD or we can assume between and calculate chalmers statistic for each. We might not know for sure the ability difference but we can quantify the potential for DIF. For each we calculate the SD across item parameters.

Something with DTF?

Do a simulation where the second dimension varies - a more real testing ground

Example with real data perhaps the Mike Frank data

Summer learning loss as applied DIF?