### 2020-02-27 paper/TEST2.Rmd

---
output: 
  pdf_document:
    keep_tex: yes
header-includes: 
- \usepackage{natbib}
- \usepackage{booktabs}
---

\begin{table}
\label{table:oos}
\caption{EDIT THIS}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

Method & Description & Literature \\

\midrule

AOAA & Test if each item has DIF by using all of the other items as anchors. Not iterative. & Originally proposed by  \\\hline

AOAA-AS & The first iteration is AOAA. All items that test positive for DIF are removed from the anchor set. Continue iterating until no new items test positive for DIF. & Proposed by \\\hline

AOAA-OAT & The first iteration is AOAA. However, only the item that shows the most extreme DIF is removed from the anchor set. Continue iterating until no new items test positive for DIF. & To our knowledge, not proposed or used previously. \\

\bottomrule
\end{tabular}
\end{table}