### 2020-02-29 paper/paper.Rmd
we're getting there!

---
output:
  bookdown::pdf_document2:
    includes:
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 5

fontsize: 11pt
bibliography: paper.bib
link-citations: yes

title: |
    New and old ways to chase a rabbit:  
    Differential item functioning detection methods
subtitle: |
    Ben Stenhaug  
    Stanford University
abstract: |
  summarize propose extensions insist as dif as student property and on communicating uncertainty
  <!-- https://tinyurl.com/ybremelq -->
header-includes: 
- \usepackage{amsmath}
- \DeclareMathOperator*{\argmax}{arg\,max} # for gini coefficient
- \DeclareMathOperator*{\argmin}{arg\,min} # for gini coefficient

- \usepackage{float}
- \usepackage[autostyle, english = american]{csquotes}

- \setlength{\parindent}{4em} # all of this for draft, papaja should take care of things
- \setlength{\parskip}{2em}
- \usepackage{setspace}\doublespacing
- \AtBeginEnvironment{tabular}{\singlespacing} # this makes that table single spaced

- \usepackage{natbib} # citations in that table were a hassle. this is still a bit broken. remember can always change tex file.
---

```{r setup, include = FALSE}
# 
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

knitr::opts_chunk$set(
    fig.align='center',
    echo=FALSE,
    warning=TRUE,
    fig.pos='H',
    fig.retina = 5
)

library(tidyverse)
library(ggtext)
library(here)
library(mirt)
library(cluster)
library(DescTools)
library(knitr)
library(kableExtra)
set.seed(123)
R.utils::sourceDirectory(here("R"))

theme_set(theme_bw())
```

\clearpage

# TO DO

- be sure all my directions match. i think i switched everything to be item easiness at some point but not sure if i did it throughout

# Introduction

```{r child = "paper_1_intro.Rmd"}
```

# DIF methods

```{r child = "paper_2a_anchor_items.Rmd"}
```

## Anchor points

```{r child = "paper_2b_anchor_points.Rmd"}
```

# Next steps

```{r child = "paper_3_next_steps.Rmd"}
```

\clearpage

# Notes

- adjust to all be easiness including multidimensional talk

# References

<!-- weird formatting that isn't apa I don't think and indentation is weird, solution is to use papaja (look at paper_apa.Rmd) or this might also be helpful https://stackoverflow.com/questions/4073674/how-to-disable-indentation-in-particular-section-in-latex-->

<!-- hopefully all of this below goes away with papaja and other solutions. but remember if it can't editing the latek file is always an option on the final draft! 
https://github.com/haozhu233/kableExtra/issues/214
https://stackoverflow.com/questions/51613152/citing-within-an-rmarkdown-table
https://tex.stackexchange.com/questions/272475/citations-in-tabular-environment-not-working
https://tex.stackexchange.com/questions/359872/is-it-possible-to-use-cite-inside-xtable
https://github.com/yihui/rmarkdown-cookbook/issues/71
-->

\bibliographystyle{apa}
\bibliography{paper.bib}