### 2020-03-10 paper/paper.Rmd
simulation went well and i think i have my final graph

---
output:
  bookdown::pdf_document2:
    includes:
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 5

fontsize: 11pt
bibliography: paper.bib
link-citations: yes

title: |
    New and old ways to chase a rabbit:  
    Differential item functioning detection methods
subtitle: |
    Ben Stenhaug  
    Stanford University
abstract: |
  summarize propose extensions insist as dif as student property and on communicating uncertainty
  <!-- https://tinyurl.com/ybremelq -->
header-includes: 
- \usepackage{amsmath}
- \DeclareMathOperator*{\argmax}{arg\,max} # for gini coefficient
- \DeclareMathOperator*{\argmin}{arg\,min} # for gini coefficient

- \usepackage{float}
- \usepackage[autostyle, english = american]{csquotes}

- \setlength{\parindent}{4em} # all of this for draft, papaja should take care of things
- \setlength{\parskip}{2em}
- \usepackage{setspace}\doublespacing
- \AtBeginEnvironment{tabular}{\singlespacing} # this makes that table single spaced

- \usepackage{natbib} # citations in that table were a hassle. this is still a bit broken. remember can always change tex file.
---

```{r setup, include = FALSE}
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

knitr::opts_chunk$set(
    fig.align='center',
    echo=FALSE,
    warning=TRUE,
    fig.pos='H',
    fig.retina = 5
)

library(tidyverse)
library(ggtext)
library(here)
library(mirt)
library(cluster)
library(DescTools)
library(knitr)
library(kableExtra)
set.seed(12345)
R.utils::sourceDirectory(here("R"))

theme_set(theme_bw())
```

<!-- the literature basically talks separately about anchor selection and DIF methods. much more literature is on DIF methods and less is on anchor selection. I'm basically writing a paper about anchor selection and just calling it DIF methods because that seems like where the action is. I'm taking fixed at that the likelihood ratio test is the way to go (and perhaps that deserves more detail)? -->

\clearpage

# Introduction {#intro}

```{r child = "paper_1_intro.Rmd"}
```

# DIF methods

```{r child = "paper_2a_anchor_items.Rmd"}
```

## Anchor points

```{r child = "paper_2b_anchor_points.Rmd"}
```

# Next steps

```{r child = "paper_3_next_steps.Rmd"}
```

\clearpage

# Notes

- adjust to all be easiness including multidimensional talk

# References

<!-- weird formatting that isn't apa I don't think and indentation is weird, solution is to use papaja (look at paper_apa.Rmd) or this might also be helpful https://stackoverflow.com/questions/4073674/how-to-disable-indentation-in-particular-section-in-latex-->

<!-- hopefully all of this below goes away with papaja and other solutions. but remember if it can't editing the latek file is always an option on the final draft! 
https://github.com/haozhu233/kableExtra/issues/214
https://stackoverflow.com/questions/51613152/citing-within-an-rmarkdown-table
https://tex.stackexchange.com/questions/272475/citations-in-tabular-environment-not-working
https://tex.stackexchange.com/questions/359872/is-it-possible-to-use-cite-inside-xtable
https://github.com/yihui/rmarkdown-cookbook/issues/71
-->

\bibliographystyle{apa}
\bibliography{paper.bib}