### 2020-03-14 paper/paper_3_sim.Rmd

To compare each of the methods in Table \ref{table:allmethods}, we conducted a simulated study. Our goal was to make the data generating process as realistic to the scenario described by @ackerman1992didactic in which some items on a math test also depend on a student’s verbal ability (the target ability is math ability, and the nuisance ability is verbal ability). As described in [the introduction](#intro), nearly all DIF simulation studies in the literature generate data by simply altering the item easiness parameters for the focal group. This setup can be re-written as a two-dimensional compensatory item response model where nuisance ability is the same for all students from the same group. One exception is @walker2017using who draw each student's target ability and nuisance ability from a two-dimensional normal distribution with varying covariance matrices.

In our simulation study, it was critical that student ability was drawn in a realistic way similar to @walker2017using. However, we don’t believe that a compensatory model is realistic in describing a math test where some items depend on verbal ability. For example, it’s hard to imagine that a student without the verbal ability to parse a word problem could fully compensate by having a higher math ability. Accordingly, we generate item responses using a simplified version of Sympson’s -@sympson1978model noncompensatory item response model in which
$$
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{2j}) = \sigma(\theta_i) \cdot \sigma(a_{2j}\eta_i)
$$
where, as before, $\theta_i$ is target ability, $\eta_i$ is nuisance ability, and $\sigma(a_{2j})$ is the item's loading on nuisance ability [@demars2016partially].

## Drawing parameters

In each run, we simulate 10,000 students with half coming from each of the reference and focal groups. For students from the reference group, target ability and nuisance ability are drawn from the two-dimensional normal distribution with mean [$\mu_\theta^\text{ref} = 0$, $\mu_\eta^\text{ref} = 0$] and covariance matrix $\begin{bmatrix} 1 & 0.5 \\ 0.5 & 1 \end{bmatrix}$. Abilities for students from the focal group are drawn using the same covariance matrix, but with means [$\mu_\theta^\text{foc} = -0.5$, $\mu_\eta^\text{foc} = -1$].

The test always has 12 items, but we vary the number of items with DIF from two to six. For items without DIF, $a_{2j} = \infty$ so that the model reduces to $\text{Pr}(y_{ij} = 1 | \theta_i) = \sigma(\theta_i)$. For items with DIF, $a_{2j}$ is calculated based on Ackerman's -@ackerman1994using angle equation as described in @walker2017using:
$$
\angle_j = \arccos \dfrac{a_{1j}^2}{a_{1j}^2 + a_{2j}^2}.
$$
An item's angle measures the relative loading of the item on the two dimensions. For example, an angle of 45$^\circ$ indicates that the item loads equally on the target ability and nuisance ability. Our simple noncompensatory model has $a_{1j} = 1$ for all items so the angle equation reduces to
$$
\angle_j = \arccos \dfrac{1}{1 + a_{2j}^2}.
$$
We are interested in specifying the angle of an item, so the relevant equation becomes
$$
a_{2j} = \sqrt{\dfrac{1 - \cos(\angle_j)^2}{\cos(\angle_j)^2}}.
$$
For DIF items, we set $a_{2j}$ based on angles with equal intervals between 20$^\circ$ and 60$^\circ$. For example, for a test with three DIF items the angles are 20$^\circ$, 40$^\circ$, and 60$^\circ$.

## Visualizing a run

```{r}
# ADD IN CODE THAT GENERATES THIS
out <- read_rds(here("paper/paper_3_out1.rds"))

sim_n_dif_items_to_foc_mean <- function(sim, n_dif_items){
    mod <-
        mod_flexible(
            sim$data,
            sim$groups,
            flex_items = (ncol(sim$data) - n_dif_items + 1):ncol(sim$data)
        )

    coef_1f(mod)$ability$b_foc[1]
}

add_foc_mean <- function(one, n_dif_items){
    one %>%
    mutate(foc_mean = sim %>% map_dbl(sim_n_dif_items_to_foc_mean, n_dif_items))
}

out <-
    out %>%
    mutate(bigsim = map2(bigsim, n_dif_items, add_foc_mean))

# this use to seem off and is a good check but seems fine now
# out %>% 
#   mutate(bigsim = bigsim %>% map(~ select(., foc_mean))) %>% 
#   unnest(bigsim) %>% 
#   ggplot(aes(x = n_dif_items, y = foc_mean)) +
#   geom_point()
```

Figure \ref{fig:difmap} provides intuition about the data generating process by showing the relationship between $\theta_i$ and $Pr(y_{ij} = 1)$ with $\eta_i$ set to the group mean for a test with six DIF items. The items are ordered by the amount of DIF such that $\angle_{j = 7} = 20^\circ$ up to $\angle_{j = 12} = 60^\circ$.

```{r difmap, fig.cap = 'For a 12 item test containing 6 items with DIF, the relationship between target ability and probability of correct response with nuisance abilities fixed to the group mean.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[5]]$pars[[1]] %>% 
  graph_pars_noncompensatory(0, -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = latex2exp::TeX("$\\theta_i$"),
    y = latex2exp::TeX("$Pr(y_{ij} = 1)$")
  )
```

Figure \ref{fig:simemmlg} shows the EM-MILG — generated using a Rasch model where both group means are fixed to 0 and item parameters are estimated freely as described in the [the anchor items section](#anchoritems) — for one run using the same item parameters that generated Figure \ref{fig:difmap}. As expected, $\tilde{d_j}$ is about $\mu_\theta^\text{foc} - \mu_\theta^\text{ref} = -0.5 - 0 = -0.5$ for the first six items which are DIF free. For the last six items, $\tilde{d_j}$ increases as $\angle_j$ increases.

```{r simemmlg, fig.cap = 'For a 20 item test, the relationship between target ability and probability of correct response with nuisance abilities fixed to group means.', out.width="70%", warning = FALSE, message = FALSE}
out$bigsim[[5]]$intuitive_mod[[1]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    labs(
          x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
      )
```

## Outcomes

For each run, we applied each AI method to find the method's identifying assumption. The method's identifying assumption was then used to fit a final model. We compared the performance of those final models according to the following outcomes.

### Achievement gap residual

An effective AI method should lead to a final model that accurately estimates the difference between the reference group's mean target ability and the focal group's mean target ability. We refer to this quantity as the achievement gap. Recall that all models set $\mu_\theta^\text{ref} = 0$, so the achievement gap reduces to $\mu_\theta^\text{ref}$. The data-generating value of $\mu_\theta^\text{foc}$ is 0.5, but each run will include sampling variability. To get at the heart of how well a method is doing, we calculated the achievement gap residual as the method's estimated achievement gap, $\hat\mu_\theta^\text{foc}$, minus the achievement gap estimated when using only the DIF-free items as anchors. In summary, this outcome measures a method’s ability to disentangle differences in target ability from nuisance ability at the group level.

### Individual abilities

Assessments are frequently used to make decisions about individual student abilities. We measured a method’s ability to do so by calculating the rank correlation between the vector of estimated abilities and the vector of true target abilities. We used rank correlation as opposed to the more typically used RMSE in agreement with Lord's -@lord1986maximum argument that RMSE can be inflated by poor performance at the extremes of ability for which the test was not designed to measure precisely. *(TODO: this isn't included in this current draft, and I'm not actually sure it's necessary. maybe this idea of rank order being better than RMSE is for another paper)*

### Anchor items.

For the methods that choose a set of anchor items, we looked directly at which anchor items were selected. An effective method should use most of the non-DIF items as anchors (the anchor hit rate) while avoiding using items with DIF as anchors (the false anchor rate).

## Results

```{r}
n_items <- 12

final_model_to_anchor_summary <- function(final_model, n_dif_items){
    coef_1f(final_model)$items %>%
        mutate(
            true_anchor =
                c(
                    rep("no dif", n_items - n_dif_items),
                    rep("some dif", ceiling(n_dif_items / 2)),
                    rep("high dif", floor(n_dif_items / 2))
                )
        ) %>%
        group_by(true_anchor) %>%
        summarize(selected = sum(anchor), total = n())
}

getfull <- function(one){
    one %>%
        select(foc_mean, ends_with("mod"), -intuitive_mod) %>%
        mutate(run = row_number()) %>%
        gather(method, model, -run, -foc_mean) %>%
        mutate(
            n_anchors = model %>% map_int(~ sum(coef_1f(.)$items$anchor)),
            est_foc_mean = model %>% map_dbl(~ coef_1f(.)$ability$b_foc[1])
        ) %>%
        filter(n_anchors != 0 | method %in% c("gini_mod", "minbc_mod")) %>% 
    mutate(resid = est_foc_mean - foc_mean)
}

what <- 
  out %>%
    mutate(achievegap = bigsim %>% map(getfull)) %>%
    select(-bigsim) %>%
    unnest(achievegap)
```

In total, we executed 100 runs for each of two, three, four, five, and six DIF items. Figure \ref{fig:achievegap} shows each method's performance on the achievement gap residual. AOAA-OAT is the clear winner. It performed nearly perfectly for two, three, or four DIF items. Even when six of the 12 items on the test contained DIF, AOAA-OAT underestimated $\mu_\theta^\text{foc}$ by only 0.05 standard deviations on its worst run. As expected, artificial DIF caused AOAA and AOAA-AS to go off the rails as the number of DIF items increases.

EMC performed better than AOAA and AOAA-AS, but worse than the other methods. MINBC and MAXGI performed similarly well with MINBC estimating the achievement gap with more precision but more bias than MAXGI, especially for tests with more than four DIF items. We hypothesize that MINBC's susceptibility to bias results from considering every item, including the items with DIF, whereas as soon as, for example, AOAA-OAT, removes an item from the anchor set, it is thereafter completely disregarded by the method.

```{r achievegap, fig.cap = 'Achievement gap residual distributions across 100 runs for each AI method and number of DIF items.', out.width="70%", warning = FALSE, message = FALSE}
what %>% 
  mutate(
    method = 
      factor(
        method,
        levels = c("AOAA_OAT_mod", "cluster_mod", "gini_mod", "AOAA_AS_mod", "AOAA_mod", "minbc_mod"),
        labels = c("AOAA-OAT", "EMC", "MAXGI", "AOAA-AS", "AOAA", "MINBC")
      )
  ) %>% 
  ggplot(aes(x = resid, y = as.factor(n_dif_items))) +
  ggridges::geom_density_ridges() +
  facet_wrap(~ method) +
  coord_flip() +
  geom_hline(yintercept = -0.5, linetype = "dashed") +
  labs(
      y = "Number of DIF items",
      x = latex2exp::TeX("Achievement gap residual")
  )
```

Figure \ref{fig:anchorfalse} shows the mean false anchor rates for each method and number of items with DIF. For example, when there were two items with DIF on the test, those two items had $\angle_{11} = 20^\circ$ and $\angle_{12} = 60^\circ$. AOAA-OAT never included item 12 in the anchor set, but incorrectly included item 11 in 60 out of the 100 runs. Accordingly, the mean false anchor rate for two DIF items and the AOAA-OAT method was
$$
\dfrac{\text{Total number of DIF items in the anchor set}}{\text{Number of DIF items on each test} \cdot \text{Number of runs}} = \dfrac{60}{2 \cdot 100} = 30\%.
$$
The fact that the item with 20$^\circ$ of DIF is most commonly incorrectly included in the anchor set is what drove the counterintuitive result that the mean false anchor rate decreases with more DIF items.

Similarly, Figure \ref{fig:anchorhit} shows the mean anchor hit rates. Remarkably, AOAA-OAT included an average of over $90\%$ of DIF-free items in the anchor set regardless of the number of DIF items on the test. Interestingly, EMC had a better anchor hit rate on tests with more DIF items. This result appears to be driven by the clustering algorithm sometimes splitting all of the DIF-free items into two separate clusters, especially when most of the items are DIF-free.

```{r anchorfalse, fig.cap = 'Mean false anchor rates across 100 runs for each AI method and number of DIF items.', out.width="70%", warning = FALSE, message = FALSE}
one_to_anchors <- function(one, n_dif_items){
    one %>%
        select(ends_with("mod"), -intuitive_mod, -gini_mod, -minbc_mod) %>%
        mutate(run = row_number()) %>%
        gather(method, model, -run) %>%
        mutate(anchor_summary = model %>% map(final_model_to_anchor_summary, n_dif_items)) %>%
        select(-run, -model) %>%
        unnest(anchor_summary) %>%
        group_by(method, true_anchor, total) %>%
        summarize(mean_selected = mean(selected)) %>%
        ungroup()
}

temp <-
    out %>%
    mutate(anchors = map2(bigsim, n_dif_items, one_to_anchors)) %>%
    select(-bigsim) %>%
    unnest(anchors)

temp %>%
    filter(true_anchor != "no dif") %>%
    group_by(n_dif_items, method) %>%
    summarize(total = sum(total), mean_selected = sum(mean_selected)) %>%
    ungroup() %>% 
    mutate(rate = mean_selected / total) %>% 
    mutate(
      method = 
        factor(
          method,
          levels = c("AOAA_OAT_mod", "cluster_mod", "AOAA_AS_mod", "AOAA_mod"),
          labels = c("AOAA-OAT", "EMC", "AOAA-AS", "AOAA")
        )
    ) %>% 
    ggplot(aes(x = n_dif_items, y = rate)) +
    geom_point() +
    facet_wrap(~ method) +
    labs(
      x = "Number of DIF items",
      y = "Mean false anchor rate"
    ) +
    scale_y_continuous(labels = scales::percent, limits = c(0, 0.5))
```

```{r anchorhit, fig.cap = 'Mean anchor hit rates across 100 runs for each AI method and number of DIF items.', out.width="70%", warning = FALSE, message = FALSE}
temp %>%
    filter(true_anchor == "no dif") %>%
    mutate(rate = mean_selected / total) %>% 
    mutate(
        method = 
          factor(
            method,
            levels = c("AOAA_OAT_mod", "cluster_mod", "AOAA_AS_mod", "AOAA_mod"),
            labels = c("AOAA-OAT", "EMC", "AOAA-AS", "AOAA")
          )
    ) %>%  
    ggplot(aes(x = n_dif_items, y = rate)) +
    geom_point() +
    facet_wrap(~ method) +
    labs(
      x = "Number of DIF items",
      y = "Mean anchor hit rate",
      color = ""
    ) +
    scale_y_continuous(labels = scales::percent, limits = c(0, 1))
```

# Discussion and summary

Measurement instruments need to be inspected for DIF so that we can be sure of the validity of the conclusions that we draw regardless of the group membership of each student. Inspired by @ackerman1992didactic, we have resurrected the DIF-as-student-property view and shown its connection to the typical way in which AI methods and DIF simulation studies have been conceptualized.

More importantly, we reviewed a variety of AI methods, proposed new AI methods, and tested their performance in a simulation study that we believe to be more realistic than the typical DIF simulation study. In particular, we simulated student ability as drawn from a two-dimensional distribution representing a student's target and nuisance ability, and then generated data using a noncompensatory item response model. Our simulation results showed that two of the most common AI methods, AOAA and AOAA-AS, perform quite poorly, especially as the number of items containing DIF grows. On the other hand, AOAA-OAT, EMC, MINGI, and MAXBC all performed reasonably.

In particular, AOAA-OAT was the clear winner, and we recommend its use whenever possible. One reason AOAA-OAT might not always be possible is that it can be computationally expensive. For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows. To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.

In addition to exploring and testing algorithmic AI methods, we introduced a method, the EM-MILG, that an analyst can use to visualize the amount of potential DIF in their data. This method can be used either to build their intuition or as a way in which they can select anchor items by hand. The EM-MILG's sibling method, the MILG, is, we believe, the best way to visualize the results of a DIF analysis after anchor items have been selected.

Future work should test these methods' performance under a greater variety of data-generating conditions. For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions. Furthermore, our work focused on the Rasch model, and it will be of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting a 2PL or 3PL item response model.